---
title: "Initial analysis on song lyrics, using #tidytex"
output: html_notebook
---


# Libraries

```{r}
library(readr)
library(tidyverse)
library(stringr)
library(tidytext)
library(lubridate)
library(wordcloud)
library(topicmodels)
library(tm)
library(stopwords)
library(quanteda)
```


# Insert data

```{r}

songs <- read_csv("./data/songs.csv")

```


# DATA CLEANSING


## Filter data, handle missing values

```{r}

# Keep songs with 10 or more characters in the lyrics
songs <- 
  songs %>% 
  filter(str_length(lyrics) > 10)
  
# Detect the language of the song
library(cld3)
songs$lang <- detect_language(songs$lyrics)

# Filter the songs that are certainly english, the genre is not missing or 
# not available and the year is 1970 or later

songs <- 
  songs %>%
  filter(lang == "en" & 
           is.na(genre) == FALSE &
           year >= 1970 &
           genre != "Not Available")
  
songs$characters <- str_count(songs$lyrics)

```



## _text cleaning

```{r}


# Create a vector with stopwords
stopwords <- c(stopwords())
  
  
# Perform cleansing & stemming in lsong lyrics text
songs$lyrics <- tolower(songs$lyrics)
songs$lyrics <- removePunctuation(songs$lyrics)
songs$lyrics <- removeNumbers(songs$lyrics)
songs$lyrics <- stripWhitespace(songs$lyrics)
songs$lyrics <- removeWords(songs$lyrics, stopwords)
songs$lyrics <- stemDocument(songs$lyrics)


saveRDS(songs, file = "data/cleaned_data.RDS")

```


# EXPLORATORY ANALYSIS

# Are there any differences between different music genres

## _How many songs per artist

```{r}
songs <- readRDS(file = "data/cleaned_data.RDS")


songs %>% 
  group_by(artist) %>% 
  count() %>% 
  filter(n > 20) %>% 
  arrange(-n)

```
* Not really informative. maybe 

## _How many songs per genre

```{r}
songs %>% 
  group_by(genre) %>% 
  count() %>% 
  arrange(-n) %>% 
  ggplot() +
  geom_col(aes(reorder(genre, -n), n))

```

* could choose specific genres for comparison. e.g. Rock, pop, hip-hop, metal, country, jazz, electronic. 

## _What kind of music uses more words

```{r}

## number of characters per song
songs %>% 
  ggplot() +
  geom_boxplot(aes(genre, characters))

```
* Hip-Hop seems to be the significantly different music genre. 

## _How many songs per genres & Artists

```{r}

songs %>% 
  group_by(genre, artist) %>% 
  count() %>% 
  arrange(genre, -n)

```
* Not informative

## _How many songs per genres & Artists

```{r}

songs %>% 
  group_by(artist, genre) %>% 
  count() %>% 
  arrange(artist, genre, -n) %>% 
  group_by(artist) %>% 
  summarise(N = n(), total_songs = sum(n)) %>% 
  arrange(-N)

```
** Possible problem as every artist belongs to just one genres. But it is known 
that a lot of artists do change their music genre through their career.

## _How many songs per Year

```{r}

songs %>% 
  group_by(year) %>% 
  count() 

```
*** There are a few songs with wrong year ... 
++

## _How many songs per Decade and Genre

```{r}

songs %>% 
  mutate(date = as_date(paste(as.character(songs$year), "-01", "-01"))) %>% 
  mutate(decade = floor_date(date, years(10))) %>% 
  group_by(decade, genre) %>% 
  summarise(N = n()) %>% 
  mutate(freq = round(N/sum(N), 2)) %>% 
  select(-N) %>% 
  spread(key = decade, value = freq) 

songs %>% 
  mutate(date = as_date(paste(as.character(songs$year), "-01", "-01"))) %>% 
  mutate(decade = floor_date(date, years(5))) %>% 
  group_by(decade, genre) %>% 
  summarise(N = n()) %>% 
  mutate(freq = round(N/sum(N), 2)) %>% 
  filter(genre %in% c("Country", "Hip-Hop", "Metal", "Pop", "Rock")) %>% 
  ggplot(aes(decade, freq, colour = genre)) +
  # geom_line() +
  geom_smooth(se = FALSE) +
  scale_y_continuous(labels = scales::percent_format()) +
  ggtitle("Smoothed frequency of total songs per Music Genre (top 5)")
  
  
  


```
** I think it would be wrong to say things about music genres. Rock is Falling - Hip hop is raising
** could make a statistical test for counts, chi-square to show differences

## _How many characters per Decade & Decade+Genre

```{r}

songs %>% 
  mutate(date = as_date(paste(as.character(songs$year), "-01", "-01"))) %>% 
  mutate(decade = floor_date(date, years(10))) %>% 
  group_by(decade) %>% 
  summarise(characters = mean(characters, na.rm = TRUE)) %>% 
  # mutate(freq = round(N/sum(N), 2)) %>% 
  # select(-N) %>% 
  spread(key = decade, value = characters)


songs %>% 
  mutate(date = as_date(paste(as.character(songs$year), "-01", "-01"))) %>% 
  mutate(decade = floor_date(date, years(10))) %>% 
  group_by(decade, genre) %>% 
  summarise(characters = round(mean(characters, na.rm = TRUE), 0)) %>% 
  # mutate(freq = round(N/sum(N), 2)) %>% 
  # select(-N) %>% 
  spread(key = decade, value = characters) 




songs %>% 
  mutate(date = as_date(paste(as.character(songs$year), "-01", "-01"))) %>% 
  mutate(decade = floor_date(date, years(10))) %>% 
  group_by(decade, genre) %>% 
  summarise(characters = round(mean(characters, na.rm = TRUE), 0)) %>% 
 ggplot(aes(decade, characters, colour = genre)) +
  geom_line() +
  # geom_smooth(se = FALSE) +
  ggtitle("How much lyrics per song for Music Genre")
  
songs %>% 
  mutate(date = as_date(paste(as.character(songs$year), "-01", "-01"))) %>% 
  mutate(decade = floor_date(date, years(10))) %>% 
  group_by(decade) %>% 
  summarise(characters = round(mean(characters, na.rm = TRUE), 0)) %>% 
 ggplot(aes(decade, characters)) +
  geom_line() +
  # geom_smooth(se = FALSE) +
  ggtitle("How much lyrics per song per 5 year interval")


songs %>% 
  group_by(genre) %>% 
  summarise(characters = round(mean(characters, na.rm = TRUE), 0)) %>% 
ggplot(aes(reorder(genre, -characters), characters)) +
  geom_col() +
  ggtitle("Average lyrics characters per song for each music genre")

```
*** It seems that Hip-hop songs contain much more lyrics (almost double). ???? 


## _Wordcloud 

```{r}
words <- 
  songs %>%
  unnest_tokens(word, lyrics) %>% 
  group_by(word) %>% 
  count() %>% 
  arrange(-n) %>% 
  head(100)

wordcloud(words = words$word, freq = words$n, max.words = 100)
```



## _Wordcloud per genre  - chech older ... comparing

```{r}
words <- 
  songs %>%
  unnest_tokens(word, lyrics) %>% 
  group_by(genre, word) %>% 
  count() %>% 
  arrange(-n) %>% 
  group_by(genre) %>% 
  top_n(n = 100, wt = n)

genres <- unique(words$genre)

for(i in 1:length(genres)){
  temp <- filter(words, genre == genres[i])
  
  # Create a word cloud
  par(bg="grey30")
  #png(file="/home/manos4/Analysis/002_Content/output/figs/WordCloud_all_posts.png",width=1000,height=700, bg="grey30")
  wordcloud(words = temp$word, freq = temp$n, col=terrain.colors(length(temp$word), alpha=0.9), random.order=FALSE, rot.per=0.3 )
  title(main =  genres[i] , font.main = 1, col.main = "cornsilk3", cex.main = 1.2)
  #dev.off()
}
  
```

*** Print top-5 genres

## _How much they repeat themselves - Rechedule code from ....




# MODELLING ON GENRES

## _Create the model

```{r}
# _Build Model ###############################################################
  

# split into words
by_word <- 
  songs %>% 
  unnest_tokens(word, lyrics)

# find document-word counts
word_counts <- 
  by_word %>%
  count(song, word, sort = TRUE) %>%
  ungroup()

word_counts


songs_dtm <- word_counts %>%
  cast_dtm(song, word, n)

songs_dtm


#songs_lda <- LDA(songs_dtm, k = 4, control = list(seed = 1234))
songs_lda <- LDA(songs_dtm, k = 3, control = list(seed = 1234))
#songs_lda_2 <- LDA(songs_dtm, k = 2, control = list(seed = 1234))


save(songs_lda, file = "objects/songs_lda_3.RDA")

```
* Create another model with 2 groups to check hip-hop & metal Vs the rest


## _Print probabilities for each song to belong to each topic

```{r}
load(file = "objects/songs_lda_3.RDA")


songs_gamma <- 
  tidy(songs_lda, matrix = "gamma") %>% 
  arrange(document, topic)

```


## _Top terms per Topic

```{r}
 # _Calculate Tables ##########################################################
  library(tidytext)
  
  ap_topics <- tidy(songs_lda, matrix = "gamma")
  
  top_terms <- 
    ap_topics %>%
    group_by(topic) %>%
    top_n(10, gamma) %>%
    ungroup() %>%
    arrange(topic, -gamma) %>% 
    rename(prob = gamma)
  
  top_terms
```


## _Are topics and genres related?

```{r}
# _Calculate Tables ##########################################################
library(tidytext)
  
ap_topics <- tidy(songs_lda, matrix = "gamma")

  
test <- right_join(ap_topics, songs[, c("song", "genre")], by = c("document" = "song"))
 

# reorder titles in order of topic 1, topic 2, etc before plotting
test %>%
  mutate(genre = reorder(genre, gamma * topic)) %>%
  ggplot(aes(factor(topic), gamma)) +
  geom_boxplot() +
  facet_wrap(~ genre)


```
*** 1 Topic are most of genres. POP, R&B, Country...
*** The 2nd topic is dominated by Hip-Hop.
*** The 3rd topic is dominated by Metal genre. although there is a significant contribution 
of Folk music


